---
title: "Cycle_Case"
author: "Rahul"
date: '2022-06-04'
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Activated required packages
```{r}
library(tidyverse)
library(janitor)
library(scales)
library(ggplot2)
library(lubridate)
rm(list = ls())
```


## Setup the directorcy and load all the files
```{r}
setwd("E:/study/no/Cyclistic_Bike_Share_Membership")
dm1 <- read.csv("202012-divvy-tripdata.csv")
dm2 <- read.csv("202101-divvy-tripdata.csv")
dm3 <- read.csv("202102-divvy-tripdata.csv")
dm4 <- read.csv("202103-divvy-tripdata.csv")
dm5 <- read.csv("202104-divvy-tripdata.csv")
dm6 <- read.csv("202105-divvy-tripdata.csv")
dm7 <- read.csv("202106-divvy-tripdata.csv")
dm8 <- read.csv("202107-divvy-tripdata.csv")
dm9 <- read.csv("202108-divvy-tripdata.csv")
dm10 <- read.csv("202109-divvy-tripdata.csv")
dm11 <- read.csv("202110-divvy-tripdata.csv")
dm12 <- read.csv("202111-divvy-tripdata.csv")
```


## Combine all the files and remove empty rows and column
```{r}
bike_rides <- rbind(dm1,dm2,dm3,dm4,dm5,dm6,dm7,dm8,dm9,dm10,dm11,dm12)
bike_rides <- janitor::remove_empty(bike_rides,which = c("rows"))
bike_rides <- janitor::remove_empty(bike_rides,which = c("cols"))
```

## Putout date from the date-time
```{r}
bike_rides$Ymd <- as.Date(bike_rides$started_at)
bike_rides$started_at <- lubridate::ymd_hms(bike_rides$started_at)
bike_rides$ended_at <- lubridate::ymd_hms(bike_rides$ended_at)
```

* Put out number of hour from time takken by rider
```{r}
bike_rides$start_hour <- lubridate::hour(bike_rides$started_at)
bike_rides$ended_hour <- lubridate::hour(bike_rides$ended_at)
```

## Calculate number of hours and minutes of ride
```{r}
bike_rides$Hours <- difftime(bike_rides$ended_at,bike_rides$started_at,units = c("hours"))
bike_rides$Minutes <- difftime(bike_rides$ended_at,bike_rides$started_at,units = c("mins"))
```

## Filter the data remove the column we don't want
```{r}
df <- bike_rides %>% filter(Hours>0)

df.null <- na.omit(df) %>% 
 select(-ride_id,-end_station_name,-end_station_id)
```

## Creating a summary data set grouped by Weekly and Count of Rides
```{r}
bikerides2 <- bike_rides %>% group_by(weekly = floor_date(Ymd,"week")
                                      ,start_hour) %>%  filter(Minutes>0) %>% 
                summarise(
                  Minutes= sum(Minutes),
                  Mean = mean(Minutes),
                  Median = median(Minutes),
                  Max = max(Minutes),
                  Min = min(Minutes),
                  Count = n()
                ) %>% ungroup()
```

## Counting hours and rides by hour
### summarize count
```{r}
bikerides2$CntMA <- forecast::ma(bikerides2$Count,28)

summary(bikerides2$Count)

xtabs(bikerides2$Count~bikerides2$start_hour)

bikerides2$Monthly <- lubridate::month(bikerides2$weekly)
```

## Plot For number of rides in a day
```{r}
bikerides2 %>% ggplot() + geom_col(aes(x=weekly, y=Count)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Count of Rides per Day", 
       subtitle = "Bases on 28 dat moving average", y = "Average rides per day")
```

## Plot for number of rides in a hour
```{r}
bikerides2 %>%  ggplot() + geom_col(aes(x=start_hour, y=Minutes))+
  scale_y_continuous(labels = comma)+
  labs(title = "Count of Rides by Hours",
       y="Rides per Hour")
```

## Creating a summary according to membership type and bike type
```{r}
bikestype<- bike_rides %>% group_by(member_casual,rideable_type,weekly=floor_date(Ymd,"week")) %>% 
  filter(Minutes>0) %>% 
  summarise(
    Minutes= sum(Minutes),
    Mean = mean(Minutes),
    Median = median(Minutes),
    Max = max(Minutes),
    Min = min(Minutes),
    Count = n()
  ) %>% ungroup()


table(bikestype$rideable_type)
```

## Plot For different types bike count 
```{r}
ggplot(bikestype) +
  geom_col(aes(x=weekly, y=Count,fill = rideable_type)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Count of Rides by Bike type" , subtitle = "For 12 months")+
  facet_wrap(~rideable_type)
```

## Plot for finding station with most number of bike started their ride
```{r}
bike_rides %>% count(start_station_name,sort = TRUE) %>% 
  top_n(21) %>% slice(2:21) %>% 
  ggplot()+ geom_col(aes(x=start_station_name, y=n))+
  labs(title = "Top 20 Start station by Ride Count",
       x = "Station Name", y="Count of Rides") +
  scale_alpha_continuous(labels=comma)+
  coord_flip()
```

## Plot for rides by different membership people
```{r}
ggplot(bikestype) +
  geom_col(aes(x=weekly, y=Count,fill = member_casual)) +
  scale_y_continuous(labels = comma) +
  labs(title = "Count of Rides by Riders" , subtitle = "Member and Casual") +
  facet_wrap(~member_casual)
```

## Plot for rides time in minutes per week
```{r}
ggplot(bikestype) + geom_col(aes(x=weekly, y=Minutes)) +
  scale_y_continuous(labels = comma) +
  facet_wrap(~rideable_type) +
  labs(title = "Total Ride Mintutes by Week")
```

## Plot for rides time(minutes) according to bike type per week
```{r}
ggplot(bikestype,aes(x=weekly, y=Minutes, fill= rideable_type)) +
  geom_col(stat = "identity",  alpha = 0.8) +
  scale_y_continuous(labels = comma) +
  labs(title = "Rides Mintues by bike Type and week", y= "Bike Trip in Minutes")+
  facet_wrap(~rideable_type)
```

